Can greedy-like heuristics be useful for solving the Weighted Orthogonal Art Gallery Problem under regular grid discretization?

### 2020-10-12 doc/paper.tex
update
Popravke na tekstu

	\section{Exact methods}
	In this section we present the exact ILP and \emph{Constraint Programming} (CP) models for solving WMSCP  under regular grid discretization, which are used  in the rest of the paper.
	\subsection{Integer linear programming model}
	Let us suppose we are given a polygon $P$ and the discretization $D(P)$ of $P$.  The task we consider is covering all points from $D(P)$ by some vertices $V=\{v_1,...,v_n\}$ of $P$ such that the sum of their weights is minimized.
	The problem is related to the known MWSC problem as follows.
	Family $\mathcal{F}$ of nonempty sets consists of the sets
	$S_i \in \mathcal{F}$ which include points $p \in D(P)$ that are visible from guard $v_i\in V$, i.e., $pv_i \subset P$.  Note that set $S_i$ includes points $p\in D(P)$ that might also be included by some other guard $v_j\in V$, $i \neq j$. For each set $S_i$, the cost $c(S_i) = w_i$ is assigned.  In this way, our starting task is equivalent of finding a minimum cardinalty covering $\mathcal{C}\subseteq\{S_1,...,S_n\}$ of the set of points $D(P)$, that is
	$$ \bigcup_{C \in \mathcal{C}} C = D(P),$$ such that $\sum_{S_i \in \mathcal{C}} c(S_i)$ is minimized.  The ILP  model for the WSCP is already known from the literature, where  here adapted to WOAGP  as follows:

	\begin{align}
	&\sum_{i=1}^n w_ix_i \longrightarrow \min  \label{eq:woagp-min}\\
	&\mbox{s.t.} \\
	&\sum_{j\in V} a_{ij}x_j \geq 1\ (\forall p_i\in D(P)) \label{eq:const-3}\\
	& x_j \in \{0,1\}, j \in V,
	\end{align}
	where
	$a_{ij} = \begin{cases}
	1, p_i \in V(j), \\
	0, \mbox{otherwise}
	\end{cases}$
	and $x_i = \begin{cases}
	1, \mbox{ if } \mbox{ the point } p_i \in \mathcal{C},\\
	0, \mbox{otherwise},
	\end{cases}$ \\

	where $V(j)$ is the set of all points from $D(P)$ that are visible from $j$-th vertex of polygon $P$.
	Set $Z = \{j \in V\mid x_i=1\}$ represents a solution of the problem w.r.t. discretization $D(P)$ of polygon $P$.
	Constraint~(\ref{eq:const-3}) enforces that any point $p_i \in D(P)$ will be visible from at least one guard from $Z$.

	In order to solve this model, we apply a general purpose solver \textsc{Cplex}~\cite{lima2010ibm}.
	\subsection{Constraint programming model} An equivalent CP model was implemented and tested by \textit{IBM ILOG CP Optimizer}~\cite{laborie2018ibm}. In this case,  Constraint~(\ref{eq:const-3}) is transformed into
	\begin{equation}
	\bigvee_{ j \in V } (a_{ij} \wedge x_j) = 1,
	\end{equation}
	whereas the other constraints and the objective function are the same like in the above ILP model. Note that CP appraoch works in a branch-and-bound manner employing a constraint propagation and variable domain filtering~\cite{rossi2006handbook}.
	\section{Algorithmic Approaches for WOAGP}
	\subsection{Greedy approaches for solving WOAGP}
 Greedy algorithms produce a solution of reasonable quality within a short interval of time and are, in essence, easy to implement. Efficiency of such  heuristic is related to a greedy criterion utilized to expand current (non-complete, i.e., partial) solution to complete one. Among all candidates (solution components for expansion, that is not-yet-considered guards) to extend current partial solution,  we choose one with the smallest greedy value and add it to the current solution until it becomes complete (i.e., it covers all  points from $D(P)$).

@@

	\begin{algorithm}[!t]
		\caption{Greedy Heuristic}\label{alg:greedy}
		\begin{algorithmic}[1]
			\State \textbf{Input:} an instance of a problem
			\State \textbf{Output:} A (feasible) non-expandable solution (or reporting that no feasible solution)
			\State $s^{ps} \gets ()$ \hspace{0.3cm}// partial solution set to empty solution
			\While{$\text{Extend}(s^{ps}) \neq \emptyset$}
			\State Select solution component $e \in  \text{Extend}(s^{ps})$ \hspace{0.3cm}//\,w.r.t.\  some criterion $g$
			\State Extend $s^{ps}$ by $e$
			\EndWhile
		\end{algorithmic}
	\end{algorithm}
	\subsubsection{An existing greedy method}

@@

	\subsubsection{A Novel Greedy Heuristic}
	In this subsection we present a novel greedy criterion. First, we introduce a term  ``incorrect point''. For a point from $D(P)$ we say that it is \textit{incorrect} if it is not covered by any guard from the current partial solution $s^{ps}$. Let us denote by $incorrect_{total}(s^{ps})$ the total number of incorrect points from discretization $D(P)$ w.r.t. $s^{ps}$. Note that  $incorrect_{total}(s^{ps}) = |D(P)|-f(s^{ps})$, where $f(s^{ps})$ is defined in (\ref{eq:sp}) and $|D(P)|$ is the cardinality of discretization set . Let $w_{total}$ be the total sum of all weights among all vertices. The greedy function w.r.t. partial solution $s^{ps}$ and  guard $v$ is defined as follows.
	\begin{equation}\label{eq:greedyfun2}
    g_2(s^{ps}, v)  =    \frac{\sum_{i \in s^{ps} \cup \{v\}} w_i}{w_{total}}+ \frac{incorrect_{total}(s^{ps}\cup v)}{|D(P)|}.
	\end{equation}

@@

	Ties occurred in the search are broken by using  price-per-unit heuristic which is stated as follows.
	For each not yet considered vertex $v_i$, we denote the region of polygon $P$ that is visible from $v_i$ by $\emph{Surf}(v_j)$. As the next candidate to extend $s^{ps}$, we choose such a guard,  with the smallest ratio between the price and the visible surface area. More precisely, this  criterion is given as:
	\begin{align}
	g'(s^{ps}, v_i) = \frac{w_{i}}{|\emph{Surf}({v_i})|},
	\end{align}

@@

	In order to enable fast calculation of greedy functions described in previous subsections, we noticed that it could be useful to allow fast updating of the number of uncovered vertices. Therefore, we introduced two useful structures: %and two auxiliary functions:
	\begin{itemize}
		%\item structure \texttt{map<Point,list<Point>> Visibility} -- for each point of $D(P)$, we provide the list of guards which are visible from that point;
		\item structure \texttt{numberOfGuards} -- which is a map structure, where each point from $D(P)$ is mapped to the number of guards in solution which cover that point;
		\item structure \texttt{CoveredPoints} -- which is a set structure which keeps the points from $D(P)$ that are covered by the partial solution;
		%\item function \texttt{updateCoveredPointsAdd (vertex v)} -- does the update of  structures %\texttt{CoveredPoints} and \texttt{numberOfGuards} by considering only new points from $D(P)$ covered by vertex $v$, when $v$  is being added to partial solution; %pseudocode of the procedure is shown in Algorithm \ref{alg:updateCoveredPointsAdd}.
		%\item function \texttt{updateCoveredPointsRemove(vertex v)} -- does the update of structures \texttt{CoveredPoints} and \texttt{numberOfGuards} by removing all points which are covered only by $v$ , when $v$ is removed from partial solution; %pseudocode  is shown in Algorithm \ref{alg:updateCoveredPointsRemove}.
	\end{itemize}
  When vertex $v$ is being added to partial solution, we update current \texttt{CoveredPoints} and \texttt{numberOfGuards} by considering only new points from $D(P)$ covered by $v$. 
  When $v$ is removed from partial solution, we update current \texttt{CoveredPoints} and \texttt{numberOfGuards} by removing all points which are covered only by $v$.  These two functions allow also a fast calculation of the functions $f$ and  $incorrect_{total}$: a candidate vertex $v$ is temporary added to the partial solution, new status is checked and then being removed from the solution.

@@

\subsubsection{A Hybrid of the \textsc{Greedy + \textsc{Cplex}}}
	The performance of \textsc{Cplex}  sooner or later degrades w.r.t. instance size due to the complexity of the problem. On the other hand, in the later stage, there is an increased chance for Greedy to worsen the obtained greedy solution. So, it makes sense to combine  partial solutions generated with a Greedy procedure over a few iterations  and only then to make use the \textsc{Cplex} to do the completion of the partial solution. In details, our approach consists of the following steps:
	\begin{enumerate}
		\item Run a Greedy method up to $K$ iterations (parameter) to obtain a partial solution $s^{ps}$ (therefore, $|s^{ps}| = K$);
		\item Take solution $s^{ps}$ and make it complete by solving a corresponding submodel via \textsc{Cplex}:
		\begin{itemize}
			\item \textsc{Cplex} solves corresponding sub-model which is formed by adding constraints $x_{i} = 1$, for all $v_i \in s^{ps}$ into the existing WOAGP model~\ref{eq:woagp-min};
			\item a complete solution $\bar {s^p}$ is obtained;
		\end{itemize}
		\item Return $f(\bar {s^p})$.
	\end{enumerate}

>>>>

	\section{Exact methods}
	In this section we present the exact ILP and \emph{Constraint Programming} (CP) models for solving WMSCP  under regular grid discretization, which are used  in the rest of the paper.
	\subsection{Integer linear programming model}
	Let us suppose we are given a polygon $P$ and the discretization $D(P)$ of $P$.  The task we consider is covering all points from $D(P)$ by some vertices $V=\{v_1,...,v_n\}$ of $P$ such that the sum of their weights is minimized.
	The problem is related to the known MWSC problem as follows.
	Family $\mathcal{F}$ of nonempty sets consists of the sets
	$S_i \in \mathcal{F}$ which include points $p \in D(P)$ that are visible from guard $v_i\in V$, i.e., $pv_i \subset P$.  For each set $S_i$, the cost $c(S_i) = w_i$ is assigned.  In this way, our starting task is equivalent of finding a minimum cardinalty covering $\mathcal{C}\subseteq\{S_1,...,S_n\}$ of the set of points $D(P)$, that is
	$$ \bigcup_{C \in \mathcal{C}} C = D(P),$$ such that

$$f(\mathcal{C}) = \sum_{S_i \in \mathcal{C}} c(S_i)$$

is minimized.

The ILP  model for the WSCP is already known from the literature, where  here adapted to WOAGP  as follows:

	\begin{align}
	&\sum_{i=1}^n w_ix_i \longrightarrow \min  \label{eq:woagp-min}\\
	&\mbox{s.t.}\nonumber \\
	&\sum_{j=1}^n a_{ij}x_j \geq 1,\ (\forall p_i\in D(P)) \label{eq:const-3}\\
	& x_j \in \{0,1\},\ j\in\{1,2,\ldots,n\}\label{eq:const-4},
	\end{align}
	where
	$a_{ij} = \begin{cases}
	1, p_i \in S_j, \\
	0, \mbox{otherwise}.
	\end{cases}$
	 \\

	Set $Z = \{v_j \in V\mid x_j=1\}$ represents a solution of the problem w.r.t. discretization $D(P)$ of polygon $P$.
	Constraint~(\ref{eq:const-3}) enforces that any point $p_i \in D(P)$ will be visible from at least one guard from $Z$.

	In order to solve this model, we apply a general purpose solver \textsc{Cplex}~\cite{lima2010ibm}.
	\subsection{Constraint programming model} An equivalent CP model was implemented and tested by \textit{IBM ILOG CP Optimizer}~\cite{laborie2018ibm}. In this case,  Constraint~(\ref{eq:const-3}) is transformed into
	\begin{equation}
	\bigvee_{ j \in V } (a_{ij} \wedge x_j) = 1,
	\end{equation}
	whereas the other constraints and the objective function are the same as in the above ILP model. Note that CP approach works in a branch-and-bound manner employing a constraint propagation and variable domain filtering~\cite{rossi2006handbook}.
	\section{Algorithmic Approaches for WOAGP}
	\subsection{Greedy approaches for solving WOAGP}
 Greedy algorithms produce a solution of reasonable quality within a short interval of time and are, in essence, easy to implement. Efficiency of such  heuristic is related to a greedy criterion utilized to expand current (non-complete, i.e., partial) solution to complete one. Among all candidates (solution components for expansion, that is not-yet-considered guards) to extend current partial solution,  we choose one with the smallest greedy value and add it to the current solution until it becomes complete (i.e., it covers all  points from $D(P)$).

@@

	\begin{algorithm}[!t]
		\caption{Greedy Heuristic}\label{alg:greedy}
		\begin{algorithmic}[1]
			\State \textbf{Input:} an instance of a problem
			\State \textbf{Output:} A (feasible) non-expandable solution $s^{ps}$  (or reporting that no feasible solution exists)
			\State $s^{ps} \gets ()$ \hspace{0.3cm}// partial solution set to empty solution
			\While{$\text{Extend}(s^{ps}) \neq \emptyset$}
			\State Select solution component $e \in  \text{Extend}(s^{ps})$  \  w.r.t.\  some criterion $g$
			\State Extend $s^{ps}$ by $e$
			\EndWhile
		\end{algorithmic}
	\end{algorithm}
	\subsubsection{An existing greedy method}

@@

	\subsubsection{A Novel Greedy Heuristic}
	In this subsection we present a novel greedy criterion. First, we introduce a term  ``incorrect point''. For a point from $D(P)$ we say that it is \textit{incorrect} if it is not covered by any guard from the current partial solution $s^{ps}$. Let us denote by $incorrect_{total}(s^{ps})$ the total number of incorrect points from discretization $D(P)$ w.r.t. $s^{ps}$. Note that 
 \begin{equation}\label{eq:incorrect}
 incorrect_{total}(s^{ps}) = |D(P)|-f(s^{ps}),
 \end{equation}

 where $f(s^{ps})$ is defined in (\ref{eq:sp}) and $|D(P)|$ is the cardinality of discretization set . Let $w_{total}$ be the total sum of all weights among all vertices. The greedy function w.r.t. partial solution $s^{ps}$ and  guard $v$ is defined as follows.
	\begin{equation}\label{eq:greedyfun2}
    g_2(s^{ps}, v)  =    \frac{\sum_{i \in s^{ps} \cup \{v\}} w_i}{w_{total}}+ \frac{incorrect_{total}(s^{ps}\cup v)}{|D(P)|}.
	\end{equation}

@@

	Ties occurred in the search are broken by using  price-per-unit heuristic which is stated as follows.
	For each not yet considered vertex $v_i$, we denote the region of polygon $P$ that is visible from $v_i$ by $\emph{Surf}(v_i)$. As the next candidate to extend $s^{ps}$, we choose such a guard,  with the smallest ratio between the price and the visible surface area. More precisely, this  criterion is given as:
	\begin{align}
	g'(s^{ps}, v_i) = \frac{w_{i}}{|\emph{Surf}({v_i})|},
	\end{align}

@@

	In order to enable fast calculation of greedy functions described in previous subsections, we noticed that it could be useful to allow fast updating of the number of uncovered vertices. Therefore, we introduced two useful structures: %and two auxiliary functions:
	\begin{itemize}
		%\item structure \texttt{map<Point,list<Point>> Visibility} -- for each point of $D(P)$, we provide the list of guards which are visible from that point;
		\item structure \texttt{numberOfGuards} -- which is a map structure, where each point from $D(P)$ is mapped to the number of guards in solution which cover that point;
		\item structure \texttt{CoveredPoints} -- as a set structure which keeps the points from $D(P)$ that are covered by the partial solution;
		%\item function \texttt{updateCoveredPointsAdd (vertex v)} -- does the update of  structures %\texttt{CoveredPoints} and \texttt{numberOfGuards} by considering only new points from $D(P)$ covered by vertex $v$, when $v$  is being added to partial solution; %pseudocode of the procedure is shown in Algorithm \ref{alg:updateCoveredPointsAdd}.
		%\item function \texttt{updateCoveredPointsRemove(vertex v)} -- does the update of structures \texttt{CoveredPoints} and \texttt{numberOfGuards} by removing all points which are covered only by $v$ , when $v$ is removed from partial solution; %pseudocode  is shown in Algorithm \ref{alg:updateCoveredPointsRemove}.
	\end{itemize}
  When vertex $v$ is being added to partial solution, we update current \texttt{CoveredPoints} and \texttt{numberOfGuards} by considering only new points from $D(P)$ covered by $v$.
  When $v$ is removed from partial solution, we update current \texttt{CoveredPoints} and \texttt{numberOfGuards} by omitting all points which are covered only by $v$.  These two functions allow also a fast calculation of  functions (\ref{eq:sp}) and  (\ref{eq:incorrect}): a candidate vertex $v$ is temporary added to the partial solution, new status is checked and then being removed from the solution.

@@

\subsubsection{A Hybrid of the \textsc{Greedy + \textsc{Cplex}}}
	The performance of \textsc{Cplex}  sooner or later degrades w.r.t. instance size due to the complexity of the problem. On the other hand, in the later stage, there is an increased chance for Greedy to worsen the obtained greedy solution. So, it makes sense to combine  partial solutions generated with a Greedy procedure over a few iterations  and only then to make use the \textsc{Cplex} to do the completion of the partial solution. In details, our approach consists of the following steps:
	\begin{enumerate}
		\item Run a Greedy method up to $K$ iterations  to obtain a partial solution $s^{ps}$, where $K$ is a parameter of the algorithm.
		\item Take solution $s^{ps}$ and make it complete by solving a corresponding submodel via \textsc{Cplex}:
		\begin{itemize}
			\item \textsc{Cplex} solves corresponding sub-model which is formed by adding constraints $x_{i} = 1$, for all $v_i \in s^{ps}$ into the existing ILP  model~(\ref{eq:woagp-min})--(\ref{eq:const-4});
			\item a complete solution $\overline s$ is obtained;
		\end{itemize}
		\item Return $f(\overline s)$.
	\end{enumerate}
